$  $% !TeX spellcheck = de_DE
%Die Angabe des schlauen Spruchs auf diesem Wege funtioniert nur,
%wenn keine Änderung des Kapitels mittels den in preambel/chapterheads.tex
%vorgeschlagenen Möglichkeiten durchgeführt wurde.
\setchapterpreamble[u]{%
	\dictum[Albert Einstein]{Probleme kann man niemals mit derselben Denkweise lösen, durch die sie entstanden sind.}
}
\chapter{A file structure for the generated code}
\label{chap: File structure}

Considering our running example for code generation in this Master thesis, the generated code is organized into the following files as explained below:

All the data types, event types, interfaces, exception types that are used in the example are stored along with the parameter channel and parameter queue as shown below:

\newpage 

\dirtree{%
	.1 src-gen.
	.2 DatatypesInterfacesEventsAndExceptions.
	.3 include.
	.4 Datatypes.h.
	.4 Exceptions.h.
	.4 FailureEvent.h.
	.4 InterfaceA.h.
	.4 InterfaceB.h.
	.4 ParameterChannel.h.
	.4 ParameterQueue.h.
}

All the constituents of the \texttt{Component\_Callee} are arranged as shown below:

\dirtree{%
	.1 src-gen.
	.2 Component\_Callee.
	.3 AutogeneratedCode.
	.4 include.
	.5 ComponentType\_Callee.h.
	.5 EventEmitterPorts\_Callee.h.h.
	.4 src.
	.5 ComponentType\_Callee.cpp.
	.5 EventEmitterPorts\_Callee.cpp.
	.3 UserCode.
	.4 include.
	.5 ComponentImplementation\_Callee.h.
	.4 src.
	.5 ComponentImplementation\_Callee.cpp.
	.2 Component\_Callee\_impl\_inst\_Instance.
	.3 AutogeneratedCode.
	.4 include.
	.5 ProvidedInterfacePorts\_Callee\_impl\_inst.h.
	.5 ComponentContainer\_Callee\_impl\_inst.h.
	.4 src.
	.5 ProvidedInterfacePorts\_Callee\_impl\_inst.cpp.
	.5 ComponentContainer\_Callee\_impl\_inst.cpp.
}

\newpage

All the constituents of the \texttt{Component\_Caller} are arranged as shown below:
\dirtree{%
	.1 src-gen.
	.2 Component\_Caller.
	.3 AutogeneratedCode.
	.4 include.
	.5 ComponentType\_Caller.h.
	.5 EventReceiverPorts\_Caller.h.
	.5 RequiredInterfacePorts\_Caller.h.
	.4 src.
	.5 ComponentType\_Caller.cpp.
	.5 EventReceiverPorts\_Caller.cpp.
	.5 RequiredInterfacePorts\_Caller.cpp.
	.3 UserCode.
	.4 include.
	.5 ComponentImplementation\_Caller.h.
	.4 src.
	.5 ComponentImplementation\_Caller.cpp.
	.2 Component\_Caller\_impl\_inst\_Instance.
	.3 AutogeneratedCode.
	.4 include.
	.5 ProvidedInterfacePorts\_Caller\_impl\_inst.h.
	.5 ComponentContainer\_Caller\_impl\_inst.h.
	.4 src.
	.5 ProvidedInterfacePorts\_Caller\_impl\_inst.cpp.
	.5 ComponentContainer\_Caller\_impl\_inst.cpp.
}